---
title: "DS_DataProd_w3 <br> A State Tax survey"
author: "Jie Yang"
date: "October 22, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE, warnings=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

## Dataset

Here, the US State goverment data for tax collection is used. The full report is at *http://rpubs.com/aceduce/DS_DataProd_w3*
Github Rmarkdown:
*https://github.com/aceduce/DSS_DataProd/blob/master/DDS_DataProd_w3.Rmd*

Key items in the survey:
State Name, Tax Item Name, Survey year. 
An example of the outputs are:

```{r comment="", echo=FALSE}
setwd("D:/Coursera/DS/Pack/DataProduct/")
df<-read.table("2016-stc-detailed.csv",stringsAsFactors=FALSE, na.strings = "X", sep=',', header = TRUE)
head(df)

#convert the numbers with characters into numerics
df$AMOUNT<-as.numeric(gsub(",","",df$AMOUNT))
```

Data source:

*https://www.census.gov/data/tables/2016/econ/stc/2016-annual.html*


## Using Plotly

This is to use plotly `library(plotly)` to do some plots and show the familarity with the `Rmarkdown`.
```{r, echo=FALSE, results='hide'}
#data exploration 
#NA portion in the tax amount
mean(is.na(df$AMOUNT))
#tax items in total, from T00 to TA5
table(df$item)

```

## Data Explorations

- For the excercise, we re-do the summation for all the data. 

```{r echo=FALSE, warning=FALSE,message = FALSE}
library("dplyr")
library("plotly")
df_sum<-df %>%
  filter(!is.na(AMOUNT) & item!="T00") %>%
  group_by(State_Name) %>%
  summarise(Total=sum(AMOUNT)) 
#compare the summations

df_s<-df[df[,"item"]=="T00", ][,c("State_Name","AMOUNT","item","ITEM_NAME")]

df_s<-df_s[order(df_s$State_Name),]
#df_s$total_add<-df_sum[,2]
df_s_states<-subset(df_s, State_Name!="United States")

# we only plot the states, not US in total.
df_sum_states<-subset(df_sum, State_Name!="United States")
p1<-plot_ly(df_sum_states, x = ~State_Name, y = ~Total, type = "bar", name="sum")%>%
  add_lines(name=~"from summations")
p2<-plot_ly(df_s_states, x = ~State_Name, y = ~AMOUNT, type = "bar",name="dataset")%>%
  add_lines(name=~"from dataset")
#Look at total contributions
p<-subplot(p1,p2)
p
```

Clearly, one can see the total is not the simple summation of different items.  I'll use the *sum from the original dataset*. 

## Contirbutions from States

We see these two states contribute most to the tax. 

```{r}
#p3<-plot_ly(df_sum_states, labels= ~State_Name, values= ~Total, type = "pie",name="sum")
p4<-plot_ly(df_s_states, labels= ~State_Name, values= ~AMOUNT, type = "pie",name="dataset", showlegend=TRUE)%>%
  add_pie(hole = 0.5) %>%
  layout(legend = ~State_Name)
  
p4
```

## California Tax in 2016

```{r}
#df_states<-df[df[,"item"]!="T00", ][,c("State_Name","AMOUNT","item","ITEM_NAME")]
df_CA<-subset(df, State_Name=="California" & item!="T00", c("State_Name","AMOUNT","item","ITEM_NAME"))

df_NY<-subset(df, State_Name=="New York" & item!="T00", c("State_Name","AMOUNT","item","ITEM_NAME"))

plot_ly(df_CA, labels= ~ITEM_NAME, values= ~AMOUNT, type = "pie")%>%
  layout(title = "California Tax in 2016")
```

## New York Tax in 2016

```{r}
plot_ly(df_NY, labels= ~ITEM_NAME, values= ~AMOUNT, type = "pie")%>%
  layout(title = "New York Tax in 2016")
```

## Comparison cont'

Let's compare the NY and CA for a bit (breakdown items)

```{r}
df_CA_NY<-rbind(df_CA,df_NY)
plot_ly(df_CA_NY, y=~ITEM_NAME, x=~AMOUNT, type="bar", color=~State_Name)
```

## Further plot

One can identify the key difference and plot against few parameters for further insights. 

- Populations
- tax rates
- Wage levels
  


## Reference

https://github.com/spujadas/coursera-ddp-plotly/blob/master/road-accidents-gb-2005-2015.Rmd

https://plot.ly/r/subplots/

https://plot.ly/r/mixed-subplots/

https://plot.ly/r/reference/



